---
title: "RSSSAM008"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Introduction
The aim of this assignment is to obtain an estimate of the location of the boat which is anchored along a piece of straight coastline at position alpha kms along the shore and a distance beta kms out to sea. The posterior of these parameters along with a given observed data set containing 200 points is used.  The Accept-Reject method allows us to find the location of the boat. This is a sampling method which allows sampling from the posterior distribution as it is difficult to simulate by an inverse transformation. 

Part a

```{r}
set.seed(1)

data <- read_excel("RandomData2021.xlsx")
x<- data[,45]

alpha = 1
beta = 1

#Making alpha and beta into one variable 
para = c(alpha,beta)

#log posterior
log_post <- function(para,x){
  alpha=para[1]
  beta =para[2]
  
 log_post2 <- (203*log(beta)-8*beta-200*(log(atan(alpha/beta)+atan((5-alpha)/beta)))+sum(log(1/(beta^2+(x-alpha)^2))))
  return(-log_post2)
}
#Finding the optimized values for alpha and beta
optim(para ,fn = log_post,x=x,method = 'L-BFGS-B' ,lower = c(0.001,0.001), upper = c(4.999,4.999))

```
The optimized values for alpha and beta are (2.3211793,0.7364848)

```{r}
log_post2 <- function(alpha,beta,x){
 log_post3 <- (203*log(beta)-8*beta-200*(log(atan(alpha/beta)+atan((5-alpha)/beta)))+sum(log(1/(beta^2+(x-alpha)^2))))
  return(exp(log_post3))
}

alpha_range <- seq(0.001,4.999,length=200)
beta_range <- seq(0.001,4.999, length =200)

(maximum<-l_post(2.3211793,0.7364848,x))

vec_log_function <- Vectorize(FUN = "log_post2", vectorize.args = c("alpha","beta"))
z<- outer(X=alpha_range,Y=beta_range,FUN = "vec_log_function", x=x)                              

#Plotting the posterior function 
contour(alpha_range,beta_range,z, nlevels= 10,
        xlab = expression(alpha),
        ylab = expression(beta),
        xlim = range(alpha_range<-2:2.9),
        ylim = range(beta_range <-0:1),
        main = "Contour of the Posterior Distribution")

#Adding a boundary box
rect(xleft= 2, ybottom = 0.4, xright=2.6, ytop = 1) 
```

The maximum value of the graph is 7.006676e-116. This was calculated by using the optim function to obtain  the optimized values for alpha and beta. These optimized values were then put into the posterior function to find the maximum value (mode). 
A bounding box was made around the contour graph, this shows the area that will be used to generate samples with the accept-reject method. 

```{r}
#Applying the AR method

ns <- 200000

alpha_range_propose <- runif(ns,2,2.6)
beta_range_propose <-runif(ns,0.4,1)
z_range_propose <-runif(ns,0,maximum)

samples <- matrix(NA,nrow=ns,ncol = 2)

for (i in 1:ns){
  if(z_range_propose[i]<= vec_log_function(alpha_range_propose[i],beta_range_propose[i],x)){
    
    samples[i,1]<- alpha_range_propose[i]
    samples[i,2]<- beta_range_propose[i]
  }
  else{next}
}

filter_samples <- na.omit(samples)
```
A sample of 200000 observations were simulated. These were then put into a uniform distribution for alpha, beta and z. Where z is the outer product of arrays of the range alpha and beta. The range of this distribution was the respective bounding box dimensions for alpha and beta, while z had a range of 0 to the maximum value of the contour graph. Points were accepted if the proposed range of alpha and beta were less than or equal to the proposed range of z and rejected otherwise. 
The red dots represent the rejection points and the green dots represent the acceptance points.

```{r}
log_post2 <- function(alpha,beta,x){
 log_post3 <- (203*log(beta)-8*beta-200*(log(atan(alpha/beta)+atan((5-alpha)/beta)))+sum(log(1/(beta^2+(x-alpha)^2))))
  return(exp(log_post3))
}

alpha_range <- seq(0.001,4.999,length=200)
beta_range <- seq(0.001,4.999, length =200)

(maximum<-l_post(2.3211793,0.7364848,x))

vec_log_function <- Vectorize(FUN = "log_post2", vectorize.args = c("alpha","beta"))
z<- outer(X=alpha_range,Y=beta_range,FUN = "vec_log_function", x=x)                              

#Plotting the posterior function 
contour(alpha_range,beta_range,z, nlevels= 10,
        xlab = expression(alpha),
        ylab = expression(beta),
        xlim = range(alpha_range<-2:2.9),
        ylim = range(beta_range <-0:1),
        main = "Contour of the Posterior Distribution")

legend("topleft", legend=c("Accept", "Reject"),
       col=c("green", "red"), lty=3:3, cex=0.8)

#Adding a boundary box
rect(xleft= 2, ybottom = 0.4, xright=2.6, ytop = 1) 

#Adding the rejection points to the plot
points(alpha_range_propose,beta_range_propose, pch=19, cex=.25, col="red")

#Adding the acceptance points to the plot
points(filter_samples, pch=19, cex=.25,col="green")


```

Part b
```{r}

#i 
#Location of the boat
alpha_boat.location <- mean(filter_samples[,1], na.rm = TRUE)
beta_boat.location <- mean(filter_samples[,2], na.rm = TRUE)

(boat_location <-c(alpha_boat.location,beta_boat.location))



#ii
#Pr(beta > 0:55)
no.beta <- length(which(filter_samples[,2]>0.55))

(prob.beta <- no.beta/length(filter_samples))

```
The location of the boat was found to be (2.2492815, 0.7483431). This was obtained by taking the mean of the accepted values for alpha and beta. 
The probability of beta being greater than 0.55 is 0.5.This was obtained by taking the amount of beta values accepted in the accept-reject method that are greater than 0.55 and dividing it by the amount of values in the entire acceptance category.


Conclusion
From this assignment it is shown that the Accept-Reject method is an effective tool for generating samples. By using this method the solution to the problem was found which was to obtain an estimate of the location of the boat which is anchored along a piece of straight coastline at position 2.32 kms along the shore and a distance 0.75 kms out to sea.


References
Textbook: Independent Random Sampling Methods
Authors : Luca Martino, David Luengo, Joaquín Míguez

